Option Returns and the risk-free Rate in Crypto Asset Markets

### 2024-03-11 written/paper.tex
progress correcting green ink

\title{Option Returns and the risk-free Rate in Crypto Asset Markets}
\author{Julian Winkel\footnote{International Research Training Group 1792;
  Humboldt-Universität zu Berlin,
  \href{mailto:julian.winkel@hu-berlin.de}{\nolinkurl{julian.winkel@hu-berlin.de}}} \and Wolfgang K. Härdle\footnote{BRC Blockchain Research Center,
  Humboldt-Universität zu Berlin, Berlin, Germany; Sim Kee Boon
  Institute, Singapore Management University, Singapore; WISE Wang Yanan
  Institute for Studies in Economics, Xiamen University, Xiamen, China;
  Dept. Information Management and Finance, National Chiao Tung
  University, Hsinchu, Taiwan, ROC; Dept. Mathematics and Physics,
  Charles University, Prague, Czech Republic, Grants -- DFG IRTG 1792
  and the Czech Science Foundation's grant no. 19-28231X / CAS XDA
  23020303 gratefully acknowledged,
  \href{mailto:haerdle@hu-berlin.de}{\nolinkurl{haerdle@hu-berlin.de}}}}
\date{18 February 2024}

@@

One of the most fundamental concepts in mathematical finance is the existence of a risk-free rate and a risky asset. 
While crypto assets are well-known for their riskiness, the existence of a risk-free rate has not been established in this new market. 
The risk-free rate in established markets derives from the practice of lending out money to low-risk, typically government institutions which owe an interest rate on top of the principal to the lender. Despite the naming convention, government issued bonds are not entirely risk-free due to the possibility of default.
Although default risk is commonly mitigated by a government's possibility to create money, the chance of a sovereign debt default is non-zero and hence not entirely risk-free.

In order to apply concepts of mathematical finance onto crypto assets, it is paramount to develop a risk-free rate - or at least a low-risk rate whose residual risk may converge to zero in time. 
An interesting candidate for a low-risk bond in the crypto asset space would be staking, which is the validation of transactions under the risk of losing a principal (the stake). Evidently, only crypto assets which follow a Proof-of-Stake (PoS) consensus mechanism instead of a Proof-of-Work (PoW) consensus mechanism can be used for staking. At the time, this is still a minor share of the crypto asset universe. We include one of those, Ether, in our experiment. However, we figure that a multitude of risks is still present in staking, such as the risk of hacks, software bugs, slashing or blockchain congestions and subsequent reduction of stakes.
Another approach to develop a risk-free rate would be lending in the DeFi space. Yet, this space may be vulnerable to similar risks as staking through hacks or software bugs in smart contracts.

@@

BTC Distribution pre filtering:
\begin{tabular}{lr}
\toprule
{} &          iv \\
\midrule
count &  5028945.00 \\
mean  &        0.90 \\
std   &        0.35 \\
min   &        0.00 \\
25\%   &        0.69 \\
50\%   &        0.83 \\
75\%   &        1.03 \\
max   &        5.00 \\
\bottomrule
\end{tabular}

BTC Distribution post filtering:
\begin{tabular}{lr}
\toprule
{} &          iv \\
\midrule
count &  4150778.00 \\
mean  &        0.86 \\
std   &        0.29 \\
min   &        0.10 \\
25\%   &        0.67 \\
50\%   &        0.80 \\
75\%   &        0.98 \\
max   &        2.50 \\
\bottomrule
\end{tabular}

ETH Distribution pre filtering:
\begin{tabular}{lr}
\toprule
{} &          iv \\
\midrule
count &  2275011.00 \\
mean  &        1.12 \\
std   &        0.42 \\
min   &        0.00 \\
25\%   &        0.87 \\
50\%   &        1.02 \\
75\%   &        1.27 \\
max   &        5.00 \\
\bottomrule
\end{tabular}

ETH post filtering:
\begin{tabular}{lr}
\toprule
{} &          iv \\
\midrule
count &  1802556.00 \\
mean  &        1.06 \\
std   &        0.34 \\
min   &        0.11 \\
25\%   &        0.84 \\
50\%   &        0.98 \\
75\%   &        1.18 \\
max   &        2.50 \\
\bottomrule
\end{tabular}

@@

\begin{minipage}{\textwidth}
    \centering
    \includegraphics[width=3.5in]{/Users/julian/src/bitcoin-vola-premium/btc/out/qq.png}
    \captionof{figure}{BTC Portfolio}
    \centering
    \includegraphics[width=3.5in]{/Users/julian/src/bitcoin-vola-premium/eth/out/qq.png}
     \captionof{figure}{ETH Portfolio}

@@

\begin{minipage}{\textwidth}
    \centering
    \includegraphics[width=3.5in]{/Users/julian/src/bitcoin-vola-premium/btc/out/vanilla/no_fees/density/moneyness>=1.05&moneyness<=1.1.png}
    \captionof{figure}{BTC KDE. 2908 Observations fall in the Interval.}
    \centering
    \includegraphics[width=3.5in]{/Users/julian/src/bitcoin-vola-premium/eth/out/vanilla/no_fees/density/moneyness>=1.05&moneyness<=1.1.png}
     \captionof{figure}{ETH KDE. 2545 Observations fall in the Interval.}

@@

\label{Kernel Density Estimator of the Batch in 1.05 and 1.1}
\href{https://github.com/QuantLet/Option-Risk-Premia/blob/master/compare.py}{\hspace*{\fill} \raisebox{-1pt}{\includegraphics[scale=0.05]{/Users/julian/src/up/spd/written/logos/qletlogo_tr.png}}\,KDE}

@@

>>>>

\title{Option Returns and the risk-free Rate in Crypto Asset Markets}
\author{Julian Winkel\footnote{International Research Training Group 1792;
  Humboldt-Universität zu Berlin,
  \href{mailto:julian.winkel@hu-berlin.de}{\nolinkurl{julian.winkel@hu-berlin.de}}} \and Wolfgang K. Härdle\footnote{BRC Blockchain Research Center,
  Humboldt-Universität zu Berlin, Berlin, Germany; Sim Kee Boon
  Institute, Singapore Management University, Singapore; WISE Wang Yanan
  Institute for Studies in Economics, Xiamen University, Xiamen, China;
  Dept. Information Management and Finance, National Yang Ming Chiao Tung
  University, Hsinchu, Taiwan, ROC; Dept. Mathematics and Physics,
  Charles University, Prague, Czech Republic; Institute for Digital Assets, Romania; Grants -- DFG IRTG 1792
  and the Czech Science Foundation's grant no. 19-28231X / CAS XDA
  23020303 gratefully acknowledged,
  \href{mailto:haerdle@hu-berlin.de}{\nolinkurl{haerdle@hu-berlin.de}}}}
\date{18 February 2024}

@@

One of the most fundamental assumptions in asset pricing and indeed a conceptional pillar of mathematical finance is the existence of a risk-free rate and a risky underlying. 
While crypto assets are well-known for their riskiness, the existence of a risk-free rate has not been established in this new market. 
The risk-free rate in established markets derives from the practice of lending out money to low-risk, typically government institutions which owe an interest rate on top of the principal to the lender. Despite the naming convention, government issued bonds are not entirely risk-free due to the possibility of default.
Although default risk is commonly mitigated by a government's possibility to create money, the chance of a sovereign debt default is non-zero and hence not entirely risk-free.

In order to apply pricing concepts of theoretical finance onto crypto assets, it is paramount to develop a risk-free rate - or at least a low-risk rate whose residual risk may converge to zero in time. 
An interesting candidate for a low-risk bond in the crypto asset space would be staking, which is the validation of transactions under the risk of losing a principal (the stake). Evidently, only crypto assets which follow a Proof-of-Stake (PoS) consensus mechanism instead of a Proof-of-Work (PoW) consensus mechanism can be used for staking. At the time, PoS-based assets represent a minor share of the crypto asset universe. We include one of those, Ether, in our experiment. However, we figure that a multitude of risks is still present in staking, such as the risk of hacks, software bugs, slashing or blockchain congestions and subsequent reduction of stakes.
Another approach to develop a risk-free rate would be lending in the DeFi space (Liu et al., 2024). Yet, this space may be vulnerable to similar risks as staking through hacks or software bugs in smart contracts.

@@

\begin{table}[ht]
\resizebox{\textwidth}{!}{\begin{tabular}{lrrrrr}
\hline
\textbf{}         & \multicolumn{1}{l}{\textbf{BTC Pre-Filtering}} & \multicolumn{1}{l}{\textbf{BTC Post-Filtering}} & \multicolumn{1}{l}{\textbf{ETH Pre-Filtering}} & \multicolumn{1}{l}{\textbf{ETH Post-Filtering}} \\
\hline
\hline
% Count             	  & 5028945    	& 4150778                         & 2275011                      & 1802556                       \\
 Mean               	& 0.90                  & 0.86                            & 1.12                           & 1.06                                              \\
 Std                   & 0.35                         & 0.29                           & 0.42                           & 0.34                                  \\
 Min                  & 0                            & 0.10                             & 0                           & 0.11                                    \\
 25\%               & 0.69                         & 0.67                            & 0.87                           & 0.84                                                 \\
 50\%                & 0.83                         & 0.80                             & 1.02                           & 0.98                                          \\
 75\%               & 1.03                         & 0.98                            & 1.27                         & 1.18                                             \\
 Max                 & 5.00                              & 2.50                             & 5.00                           & 2.50                               
\end{tabular}}
\caption{IV Summary Statistics Pre and Post Filtering}
\label{btc_eth_iv_summary}
\end{table}

%
%
%ETH Distribution pre filtering:
%\begin{tabular}{lr}
%\toprule
%{} &          iv \\
%\midrule
%count &  2275011.00 \\
%mean  &        1.12 \\
%std   &        0.42 \\
%min   &        0.00 \\
%25\%   &        0.87 \\
%50\%   &        1.02 \\
%75\%   &        1.27 \\
%max   &        5.00 \\
%\bottomrule
%\end{tabular}
%
%ETH post filtering:
%\begin{tabular}{lr}
%\toprule
%{} &          iv \\
%\midrule
%count &  1802556.00 \\
%mean  &        1.06 \\
%std   &        0.34 \\
%min   &        0.11 \\
%25\%   &        0.84 \\
%50\%   &        0.98 \\
%75\%   &        1.18 \\
%max   &        2.50 \\
%\bottomrule
%\end{tabular}
%
%BTC Distribution pre filtering:
%\begin{tabular}{lr}
%\toprule
%{} &          iv \\
%\midrule
%count &  5028945.00 \\
%mean  &        0.90 \\
%std   &        0.35 \\
%min   &        0.00 \\
%25\%   &        0.69 \\
%50\%   &        0.83 \\
%75\%   &        1.03 \\
%max   &        5.00 \\
%\bottomrule
%\end{tabular}
%
%BTC Distribution post filtering:
%\begin{tabular}{lr}
%\toprule
%{} &          iv \\
%\midrule
%count &  4150778.00 \\
%mean  &        0.86 \\
%std   &        0.29 \\
%min   &        0.10 \\
%25\%   &        0.67 \\
%50\%   &        0.80 \\
%75\%   &        0.98 \\
%max   &        2.50 \\
%\bottomrule
%\end{tabular}

@@

\begin{minipage}{\textwidth}
    \centering
    \includegraphics[width=3.5in]{/Users/julian/src/bitcoin-vola-premium/btc/out/qq.png}
    \captionof{figure}{QQ Plot of beta-neutral daily \textbf{BTC} Returns vs. $t$-Distribution}
    \centering
    \includegraphics[width=3.5in]{/Users/julian/src/bitcoin-vola-premium/eth/out/qq.png}
     \captionof{figure}{QQ Plot of beta-neutral daily \textbf{ETH} Returns vs. $t$-Distribution}

@@

\begin{minipage}{\textwidth}
    \centering
    \includegraphics[width=5in]{/Users/julian/src/bitcoin-vola-premium/btc/out/vanilla/no_fees/density/moneyness>=1.05&moneyness<=1.1.png}
    \captionof{figure}{Kernel Density Estimation over 2908 BTC Portfolio Returns for Moneyness in $[1.05, 1.1]$. }
    \centering
    \includegraphics[width=5in]{/Users/julian/src/bitcoin-vola-premium/eth/out/vanilla/no_fees/density/moneyness>=1.05&moneyness<=1.1.png}
     \captionof{figure}{Kernel Density Estimation over 2545 ETH Portfolio Returns for Moneyness in $[1.05, 1.1]$}

@@

\href{https://github.com/QuantLet/Option-Risk-Premia/blob/master/compare.py}{\hspace*{\fill} \raisebox{-1pt}{\includegraphics[scale=0.05]{/Users/julian/src/up/spd/written/logos/qletlogo_tr.png}}\,KDE}

@@

\hypertarget{refs}{}
\leavevmode\hypertarget{liu}{}%
Liu, F., Packham, N., Sepp, A. and Härdle WK (2024). Jump risk premia in the presence of clustered jumps. Available at SSRN 4735365.
